---
title: "STAT 240 Discussion 4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
library(tidyverse)
library(lubridate)
source("../../scripts/viridis.R")
Sys.setlocale("LC_TIME", "English_United States.1252") 
x <- seq(Sys.Date() - 7, Sys.Date(), by = 1)
Sys.setlocale("LC_TIME", "C") #since I'm at a non-English locale
factor(weekdays(x, TRUE), 
          levels = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"),
          labels = c("Mon", "Tues", "Wed", "Thurs", "Fri", "Sat", "Sun"))
```

## Group XXX (use the assigned group name/number)

## Members Present

- ADD THE NAMES OF EVERYONE PRESENT
- MAKE THE NAME **BOLD** OF THE RESPONSIBLE PERSON TO UPLOAD THE FILES FOR THE GROUP

## Members Absent

- ADD NAMES OF ANY ABSENT MEMBERS

## Questions

### 1

> The code block below reads in the data, and then modifies it. Add comments by replacing the text XXX to briefly summarize what the following line of code does.

```{r atlanta}
### Input a file where the delimiter character is a XXX
atl = read_tsv("../../data/atlanta-2018.tsv") %>%
  ## eliminates the column that ends with "wait"
  select(-ends_with("wait")) %>%
  ## eliminates the column that contains the letter "wait"
  select(-contains("_n_")) %>%
  ## XXX
  select(-terminal,-hour,-all_excluded,-all_booths) %>%
  ## XXX
  rename(passengers = all_total,
         flights = all_flights) %>%
  group_by(airport,date) %>%
  ## XXX
  summarize(passengers = sum(passengers),
            flights = sum(flights)) %>%
  ## XXX
  mutate(date = mdy(date)) %>%
  ## XXX
  mutate(wday = wday(date, label=TRUE)) %>%
  ## XXX
  mutate(month = month(date, label=TRUE)) %>%
  ## XXX
  select(airport,date,month,wday,everything()) %>% 
  ## XXX
  arrange(date)
```

### 2

>   After problem 1, the data frame `atl` should have one row per day of the year and include variables passengers and flights for the totals of each day. Create a summary data frame named `atl_sum` with columns for the average number of international flight arrivals per day and the average total number of daily passengers arriving on international flights per day in Atlanta by day of the week.
This data frame will have three columns, one for day of the week, and one for each of the two quantiative variable.

```{r}
atl_sum <- atl %>%
  group_by(wday) %>%
  summarize(passengers=mean(passengers), flights = mean(flights))
atl_sum
```

### A

> Display the summary data frame in a table with the rows arranged in order by days of the week from Sunday through Saturday.

```{r}
atl_sum
```

### B

> Display the summary data frame in a table with the rows arranged in order from most to least average number of international flight arrivals.

```{r}
atl_sum %>%
  arrange(by=desc(flights))
```

### 3

> On what single date in 2018 did the largest number of international passengers arrive at the Atlanta airport? How many passengers were there? What day of the week was this date?

```{r}
atl %>%
  filter(str_detect(date, '2018-')) %>%
  slice_max(passengers, n=1)
```

REPLACE THIS TEXT WITH YOUR RESPONSE

### 4

> Read in the data for Chicago using the appropriate `read_*()` command. Make a bar plot that shows the mean number of passengers arriving per day to O'Hare international airport by day of the week. (Each row in the data set represents an hour of a day. You will need to transform the data to obtain the values to graph. Borrow code from problem 1 to help.)

```{r}
ord = read_csv2("../../data/chicago-2018.csv") %>%
  select(-ends_with("wait")) %>%
  select(-contains("_n_")) %>%
  rename(passengers = all_total,
         flights = all_flights) %>%
  group_by(airport,date) %>%
  summarize(passengers = sum(passengers),
            flights = sum(flights)) %>%
  mutate(date = mdy(date)) %>%
  mutate(wday = wday(date, label=TRUE)) %>%
  mutate(month = month(date, label=TRUE)) %>%
  select(airport,date,month,wday,everything()) %>% 
  arrange(date)

ord1 <- ord %>%
  group_by(wday) %>%
  summarize(passengers=mean(passengers))

ggplot(ord1, aes(x=wday, y=passengers)) +
  geom_col(fill="blue") +
  xlab("Day") +
  ylab("Passengers") +
  ggtitle("Number of Passengers Per Day Arriving Chicago ORD")

```

### 5

> Read in the data for Denver using the appropriate `read_*()` command. Make a line plot that shows the total number of international flights arriving to Denver international airport by month. (Note: add `group = 1` as an aesthetic so that there is a single line for all months.)

```{r}
den = read_csv("../../data/denver-2018.csv") %>%
  ## removes cols whose names end with 'wait'
  select(-ends_with("wait")) %>%
  ## removes cols whose names contain '_n_'
  select(-contains("_n_")) %>%
  ## removes cols terminal, hour, all_excluded, all_booths
  select(-terminal,-hour,-all_excluded,-all_booths) %>%
  ## renames col all_total to passengers and col all_flights to flights
  rename(passengers = all_total,
         flights = all_flights) %>%
  group_by(airport,date) %>%
  ## modify passenger and flight cols so they show the sums of the respective categories
  summarize(passengers = sum(passengers),
            flights = sum(flights)) %>%
  ## modifies date to be in month-day-year form
  mutate(date = mdy(date)) %>%
  ## add new wday col to display weekday in string form
  mutate(wday = wday(date, label=TRUE)) %>%
  ## add new month col to display month in string form
  mutate(month = month(date, label=TRUE)) %>%
  ## reorders the cols as below
  select(airport,date,month,wday,everything()) %>% 
  ## orders cols in asc date order
  arrange(date)

den_data <- den %>% 
  select(month, flights) %>%
    group_by(month) %>%
  summarize(flights = sum(flights))

ggplot(den_data,aes(x=month, y=flights)) +
  geom_line(aes(group = 1)) +
  xlab("Months") +
  ylab("Flights") +

```

### Lubridate practice

Code below creates a datetime object with the current date and time.
Write one line of code to extract from this datetime the requested information.
(Your answers will depend on the exact time you knit the document.)

```{r}
now_dt = now()
```

### 6

> The year

```{r}
year(now_dt)
```

### 7

> The month as a number

```{r}
month(now_dt)
```

### 8

> The day of the week as a number

```{r}
wday(now_dt)
```

### 9

> The day of the week as a label

```{r}
wday(now_dt, label=TRUE)
```

### 10

> The hour

```{r}
hour(now_dt)
```
