---
title: "STAT 240 Discussion 8"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
library(tidyverse)
source("../../scripts/viridis.R")
source("../../scripts/ggprob.R")
```

## Group XXX 

## Members Present

- ADD THE NAMES OF EVERYONE PRESENT
- MAKE THE NAME **BOLD** OF THE RESPONSIBLE PERSON TO UPLOAD THE FILES FOR THE GROUP

## Members Absent

- ADD NAMES OF ANY ABSENT MEMBERS


## Questions

### 1

For each scenario, explain why the random variable does *not* have a binomial distribution.

1. A bucket contains 10 colored tokens with five that are red and five that are blue. Four tokens are drawn at random from the bucket one at a time, but without replacing the tokens drawn. $X_1$ is the number of red tokens selected.

REPLACE THIS TEXT WITH YOUR RESPONSE

2. A fair coin is tossed repeatedly until the tenth head is tossed. $X_2$ is the number of tails tossed prior to the tenth head.

REPLACE THIS TEXT WITH YOUR RESPONSE

3. Four buckets each contain a total of five tokens each, some red and some blue. The number of red tokens in the buckets are 1, 2, 3, and 4 with blue tokens making up the remainder. One token is drawn at random from each bucket. $X_3$ is the total number of red tokens drawn.

REPLACE THIS TEXT WITH YOUR RESPONSE

### Information about `dbinom()`

The function `dbinom(x, size, prob)` calculates binomial probabilities.

- `x` is the possible outcome
- `size` is the parameter $n$ for the total number of trials
- `prob` is the single trial success probability $p$.

The function is *vectorized* which means that each argument can be replaced by a vector of values instead of a single value.
The length of the vector of the output will be whichever input vector is longest. Other values are repeated cyclically, if necessary.
This will be easiest to use if we set two of the arguments to single values and only set one to be a vector of more than one value, but at times we will use vectors for more than one argument.

##### Examples

> $P(X = x)$ for $n=5$, $p=0.4$ and $x=0,\ldots,5$.

```{r}
dbinom(0:5, 5, 0.4)
```

> $P(X = 2)$ for $p = 0.4$ and $n = 1, 2, 4, 8, 16$.

```{r}
dbinom(2, 2^(0:4), 0.4)
```

> $P(X = 2)$ for $n=5$ and $p = 0.1, 0.3, 0.5, 0.7, 0.9$.

```{r}
dbinom(2, 5, seq(0.1, 0.9, 0.2))
```

Other functions such as `pbinom()` and `qbinom()` are also vectorized.

### 2

Create a data frame with columns `n`, `p`, `x`, `mu`, and `prob` where:

- `n` varies from 2 to 100 by twos (so each `n` value is even);
- `p` equals 0.5;
- `x` is $n/2$;
- `mu` is the mean of the distribution;
- `prob` is $P(X = x)$

Show the first few rows using `head()`.

```{r}
df <- data.frame(n = seq(2, 100, by=2), p = 0.5) %>%
  mutate(x=)
df
```

**(a)** What is the relationship between `x` and `mu`?

REPLACE THIS TEXT WITH YOUR RESPONSE

**(b)** Make a line graph of `prob` versus `n`.

```{r}

```

**(c)** Describe the pattern: how does the probability that a random variable is equal to the mean (when the mean is a possible value) change as $n$ increases?

REPLACE THIS TEXT WITH YOUR RESPONSE



### 3

The central limit theorem implies that the binomial distribution converges to a normal distribution as $n$ increases.
This problem will examine one aspect of the convergence, namely the right tail probability of being more than two standard deviations above the mean,
$P(X > \mu + 2\sigma)$.

**(a)** What is the probability that a normal random variable with mean $\mu$ and standard deviation $\sigma$ exceeds $\mu + 2 \sigma$? Display answer rounded to four decimal places.

```{r}

```

REPLACE THIS TEXT WITH YOUR RESPONSE


**(b)** 

Create a data frame with columns `n`, `p`, `mu`, `sigma`, `x`, and `prob` where:  
- `n` varies from 1 to 1000 by ones;  
- `p` equals 0.5;  
- `mu` is the mean of the distribution;  
- `sigma` is the standard deviation of the distribution;  
- `x` equals `mu` + 2*`sigma`;  
- `prob` is $P(X > x)$  

Display the first few rows of the data frame with `head()`.

```{r}

```

**(c)** Plot the right tail probabilities versus $n$. Add a red, dashed, horizontal line at the value you found in part **(a)**.
Add a smooth curve which lessens the visual impact of the oscillations due to the discreteness of the binomial distribution.

```{r}

```

**(d)** Repeat parts **(b)** and **(c)** if $p = 0.005$.

```{r}

```

```{r}

```

**(e)**

Ignoring the oscillations, how do the patterns of the two smooth curves in the graphs differ from one another, especially when $n$ is large?

REPLACE THIS TEXT WITH YOUR RESPONSE

### 4

Draw graphs of the binomial distributions for $n=500$ and $p = 0.5$ and $p = 0.005$,
scaled so that the x axis is restricted to where the probabilities are relatively large.  (One graph for each of the two $p$'s.)
Overlay each plot with a red normal density with a mean and standard deviation that matches the mean and standard deviation of the corresponding binomial distribution.
(Use functions in *gprob.R* for these graphs and set `scale=TRUE`.)

Compare the skewness of the distributions. Comment on how this might help explain the differences in the right tail probabilities from Problem 3.

```{r}

```

REPLACE THIS TEXT WITH YOUR RESPONSE
