---
author: "HYUN KO"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
library(tidyverse)
library(lubridate)
```

## Assignment 2

```{r read-data, echo = FALSE}
## The echo = FALSE argument to this chunk
##   means that the code will not appear in the outut
##   but the code will run so that the data is in the session

## Read Lake Monona data
## Remove rows with missing data
monona = read_csv("../../data/lake-monona-clean-2020.csv") %>%
  drop_na()
```

### Problems

### 1

The following code makes a histogram of the `days` variable in the Lake Monona data set.

```{r problem1}
ggplot(monona, aes(x=days)) +
  geom_histogram(binwidth=10,color="black",fill="white")
```

What is the approximate length in days during a typical winter over the past 170 years or so that Lake Monona has been at least 50% covered with ice?

### Response

Approximately 100 Days

### 2

Modify the code below so that:

- one of the bars is centered at 100 days
- the width of each bar is 7 days
- the fill color is your favorite color
- the x label says "Days Closed with Ice"
- the y label says "total"
- there is a title with words of your choosing

```{r problem2}
ggplot(monona, aes(x=days)) +
  geom_histogram(center=100, binwidth=7, color="black", fill="white") +
  xlab("Days Closed with Ice") +
  ylab("total") +
  ggtitle("Lake Monona Days Closed by Ice (50% or more)")
```

### 3

Code in the next chunk makes a line plot that shows how the variable `days` changes with time (using `year1`).

```{r problem3}
ggplot(monona, aes(x=year1, y=days)) +
  geom_line() +
  geom_hline(yintercept=0)
```

What does the line of code `geom_hline(yintercept=0)` do?

### Response
Indicating the minimum days of freezing is zero.


### 4

Modify the code in the following chunk so that:

- There are more descriptive axis labels
- There is an informative plot title
- There is a smooth trend line that has is colored red

```{r problem4}
ggplot(monona, aes(x=year1, y=days)) +
  geom_line() +
  geom_hline(yintercept=0) +
  geom_smooth(method="lm", color="red", se=FALSE) +
  xlab("Year") +
  ylab("Days Frozen") +
  ggtitle("Number of Days Frozen by Year")
```

- Describe the pattern of the trend.
- How long was Lake Monona closed with ice in a typical year near 1850?
- How long was Lake Monona closed with ice in a typical year near the present?
- About how many days less is the period of closure with ice changing per decade?
- Has the change been fairly steady, or has the change been substantially faster or slower in some time periods?

### Response

1. Steadily decreasing in days of freezing
2. Roughly 130 Days
3. Roughly 80 Days
4. 2 Days lesser every dacade
5. Fairly Steady
